\thetitle

### 2008-05-29 paper.tex
Tweaking...

The simplifying assumptions made by most traffic models can be classified into three categories. We use the terms \textit{uniformity}, \textit{marginality}, and \textit{regularity} do describe these classes of assumptions. They are all assumptions about the ``sameness'' of aspects of network behaviors; they differ, however, in whether the make stipulations about uniformity within each flow, across all flows, or within groups of flows. In the following sections, we explore how various modeling concepts can be expressed in the framework of the General Matrix Model. Models are expressed as transformations that will convert arbitrary traffic matrix instances into similar ones that satisfy the model's assumptions. These transformations will later allow us to evaluate the effect of each model's simplifications on the realism of generated workloads.

@@

\begin{table*}
\begin{center}
\small
\begin{tabular}{|c|c|l|}
\multicolumn{1}{c}{\textbf{model}} &
\multicolumn{1}{c}{\textbf{transformation}} &
\multicolumn{1}{c}{\textbf{description}} \\
\hline
{\footnotesize{GMM}} & --- & the general matrix model \\\hline\hline
time uniform & $\mat{A}'=\mat{A}(\ones{d_t \times d_t}/d_t)$ & all start times are equally likely \\\hline
node uniform & $\mat{S}'=\mat{S}(\ones{n \times n}/n)$~~~~$\mat{D}'=\mat{D}(\ones{n \times n}/n)$ & nodes behave stochastically identically with respect to flows \\\hline
full uniform & \textit{all uniform transformations} & applies all the uniformity assumptions of the above three models \\\hline\hline
time marginal & $\mat{A}'=(\ones{f \times f}/f)\mat{A}$ & start times chosen from marginal distribution of start times \\\hline
size marginal & $\mat{B}'=(\ones{f \times f}/f)\mat{B}$ & flow size (bytes) is chosen independently of other properties \\\hline
node marginal & $\mat{S}'=(\ones{f \times f}/f)\mat{S}$~~~~$\mat{D}'=(\ones{f \times f}/f)\mat{D}$ & nodes behave distinctly, but behavior is independent of flows \\\hline
packet marginal & $\mat{Z}'=(\ones{f \times f}/f)\mat{Z}$~~~~$\mat{V}'=(\ones{f \times f}/f)\mat{V}$ & packet behavior is variable bit-rate, aggregated across all flows \\\hline
full marginal & \textit{all marginal transformations} & applies all the marginality assumptions of the above four models \\\hline\hline
source regular & $\mat{F}'=(\trans{\mat{SS}}/f)\mat{F}$ & flow behaviors are aggregated on a per source-node basis \\\hline
dest regular & $\mat{F}'=(\trans{\mat{DD}}/f)\mat{F}$ & flow behaviors are aggregated on a per destination-node basis \\\hline
mixed regular & $\mat{F}'=\frac{1}{2f}(\trans{\mat{SS}}+\trans{\mat{DD}})\mat{F}$ & flow behaviors are aggregated on a mixed source/destination basis \\\hline
\end{tabular}
\caption{Matrix-based traffic models evaluated by paired differential simulation.}
\label{tab:traffic-models}
\end{center}
\vspace{-2em}
\end{table*}

@@

The 24-hour trace is split into 144 10-minute segments, each of which serves as the basis for a set of simulations using different traffic models. The traffic models range from a completely realistic trace-driven model, to a standard \caps{CBR} traffic model. Various partially synthetic intermediate models, described in \Section{traffic-models}, are simulated to study the impact of different aspects of traffic behavior on network performance. To preserve the fairness of the performance comparison, we keep as many features as possible constant across different traffic models. The traffic generated by each synthetic model preserves as many characteristics from the original trace as possible, within the constraints of the model. Moreover, the following features are preserved across all models: the numbers of wireless nodes, the number of flows, the number of application-initiated data units sent, the total bytes of application data sent, and the average flow duration (and therefore the average data rate).

@@

The simulations described in \Section{simulations} provide us with the raw data to compare performance metrics for synthetic traffic models with those for real traffic traces. To assess the realism of these models, however, we need a measure of how inaccurate the synthetic performance values are when compared to the real values. Let $x$ be the value of a performance metric using real traffic, and $y$ the value of the same metric using an alternate traffic model, $\M$. Some common measures of error are the difference [$y-x$], the ratio [$y/x$], and the standard error [$(y-x)/x$]. These are all reasonable measures of error; but which is most appropriate for assessing the realism of performance metrics? Instead of picking one arbitrarily, we will first consider the properties that an ideal error function should have, and then use those properties to determine the best measure of error. Moreover, we will show that the unique measure of error that exhibits these ideal properties is the  \textit{log-ratio} of metric values:
\begin{align}
\log(y/x)=\log(y)-\log(x).
\end{align}
In this discussion, $E(x,y)$ is a generic error function applied to the synthetic value, $y$, with respect to the real value, $x$.

The first property that an error function should have is  \textit{insensitivity to common factors}. That is, if both values are scaled by the same constant, the error should be unaffected:
\begin{align}
\label{eqn:normalization-invariance}
\forall\: x,y,c: E(x c,y c) &= E(x,y)
\end{align}
There are three major motivations for this requirement:
\begin{enumerate}
\item Changing units should not affect error values.
\item Error values for ``large'' and ``small'' scenarios should be directly comparable. Scenarios with large $x$ values will naturally have larger raw differences between $x$ and $y$. This requirement allows scenarios of different scales to be compared fairly and without bias.
\item Changing between metrics that differ by a known constant for each scenario should not affect error values.
\end{enumerate}
The last point is best illustrated by an example. Consider two closely related performance metrics: average throughput, $t$, and total bytes received, $r$. Suppose that there are $f$ flows in a given scenario with average duration, $d$. Since $t =r/fd$, the metrics $t$ and $r$ contain the same information---they differ only by a known constant in each simulation scenario. Equation~\ref{eqn:normalization-invariance} ensures that the errors of these metrics are the same:
\begin{align}
E(t_{\text{trc}},t_{\text{mod}})
	= E\left({\frac{r_{\text{trc}}}{fd},\frac{r_{\text{mod}}}{fd}}\right)
	= E(r_{\text{trc}},r_{\text{mod}}).
\end{align}
The difference measure does not satisfy Equation~\ref{eqn:normalization-invariance}, but the ratio, standard error, and log-ratio error measures all do.

@@

It is easily verified that the difference, ratio, and standard error measures do not satisfy \Equation{linear-compounding}, and the difference, as noted, does not satisfy \Equation{normalization-invariance}. The log-ratio is the only metric presented that satisfies both conditions. Moreover, it can be proved that $\log(y/x)$ is the \textit{only} differentiable function that satisfies both (up to a constant). In the Appendix, we present a proof of this claim. Throughout the rest of the paper, we use the log-ratio to measure the error of performance metrics.
% TODO: should Appendix be capitalized?
% TODO: hyperref the Appendix for PDF.

@@

\begin{sidewaysfigure*}
\subfloat[\textbf{Application:} Average End-to-End Delay]%
{\includegraphics[width=4.45in]{plots/Application/delay.pdf}}%
\subfloat[\textbf{Application:} Packet Delivery Ratio]%
{\includegraphics[width=4.45in]{plots/Application/delivery_ratio.pdf}}
%
\subfloat[\textbf{Application:} Average Received Throughput]%
{\includegraphics[width=4.45in]{plots/Application/throughput.pdf}}%
\subfloat[\textbf{Network:} AODV Control Overhead]%
{\includegraphics[width=4.45in]{plots/Network/control_packets.pdf}}
%
\subfloat[\textbf{Network:} Packets Dropped in Routing Queue]%
{\includegraphics[width=4.45in]{plots/Network/dropped.pdf}}%
\subfloat[\textbf{Link:} 802.11 Control Overhead]%
{\includegraphics[width=4.45in]{plots/Link/overhead.pdf}}%
%
\caption{\footnotesize%
Box-and-whisker plots of log-ratio error values for all metrics and traffic models. The lower axis indicates the log-ratio, while the upper axis shows raw ratio values. Each box contains the central majority of log-ratio values: the left and right bounds are at the $25^\text{th}$ and $75^\text{th}$ percentiles. The dark middle line indicates the median value, while the diamond marks the mean. The whiskers (dotted lines) extend to the furthest non-outlier values, while the points beyond that are outliers. The notch in the middle of each bar indicates a $95\%$ confidence interval for the true underlying median value; if two notches do not overlap, they are very unlikely to have the same median.
}
\label{fig:box-plots}
\end{sidewaysfigure*}

>>>>

The simplifying assumptions made by most traffic models can be classified into three categories. We use the terms \textit{uniformity}, \textit{marginality}, and \textit{regularity} to describe these classes of assumptions. They are all assumptions about the ``sameness'' of aspects of network behaviors; they differ, however, in whether the make stipulations about uniformity within each flow, across all flows, or within groups of flows. In the following sections, we explore how various modeling concepts can be expressed in the framework of the General Matrix Model. Models are expressed as transformations that will convert arbitrary traffic matrix instances into similar ones that satisfy the model's assumptions. These transformations will later allow us to evaluate the effect of each model's simplifications on the realism of generated workloads.

@@

\begin{table*}
\begin{center}
\small
\begin{tabular}{|c|c|l|}
\multicolumn{1}{c}{\textbf{model}} &
\multicolumn{1}{c}{\textbf{transformation}} &
\multicolumn{1}{c}{\textbf{description}} \\
\hline
{\footnotesize{GMM}} & --- & the general matrix model \\\hline\hline
time uniform & $\mat{A}'=\mat{A}(\ones{d_t \times d_t}/d_t)$ & all start times are equally likely \\\hline
node uniform & $\mat{S}'=\mat{S}(\ones{n \times n}/n)$~~~~$\mat{D}'=\mat{D}(\ones{n \times n}/n)$ & nodes behave stochastically identically with respect to flows \\\hline
full uniform & \textit{all uniform transformations} & applies all the uniformity assumptions of the above models \\\hline\hline
time marginal & $\mat{A}'=(\ones{f \times f}/f)\mat{A}$ & start times chosen from marginal distribution of start times \\\hline
size marginal & $\mat{B}'=(\ones{f \times f}/f)\mat{B}$ & flow size (bytes) is chosen independently of other properties \\\hline
node marginal & $\mat{S}'=(\ones{f \times f}/f)\mat{S}$~~~~$\mat{D}'=(\ones{f \times f}/f)\mat{D}$ & nodes behave distinctly, but behavior is independent of flows \\\hline
packet marginal & $\mat{Z}'=(\ones{f \times f}/f)\mat{Z}$~~~~$\mat{V}'=(\ones{f \times f}/f)\mat{V}$ & packet behavior is variable bit-rate, aggregated across all flows \\\hline
full marginal & \textit{all marginal transformations} & applies all the marginality assumptions of the above four models \\\hline\hline
source regular & $\mat{F}'=(\trans{\mat{SS}}/f)\mat{F}$ & flow behaviors are aggregated on a per source-node basis \\\hline
dest regular & $\mat{F}'=(\trans{\mat{DD}}/f)\mat{F}$ & flow behaviors are aggregated on a per destination-node basis \\\hline
mixed regular & $\mat{F}'=\frac{1}{2f}(\trans{\mat{SS}}+\trans{\mat{DD}})\mat{F}$ & flow behaviors are aggregated on a mixed source/destination basis \\\hline
\end{tabular}
\caption{Matrix-based traffic models evaluated by paired differential simulation.}
\label{tab:traffic-models}
\end{center}
\vspace{-2em}
\end{table*}

@@

The 24-hour trace is split into 144 10-minute segments, each of which serves as the basis for a set of simulations using different traffic models. The traffic models range from a completely realistic trace-driven model, to a standard \caps{CBR} traffic model. Various partially synthetic intermediate models, described in \Section{traffic-models}, are simulated to study the impact of different aspects of traffic behavior on network performance. To preserve the fairness of the performance comparison, we keep as many features as possible constant across different traffic models. The traffic generated by each synthetic model preserves as many characteristics from the original trace as possible, within the constraints of the model. Moreover, the following features are preserved across all models: the number of wireless nodes, the number of flows, the number of application-initiated data units sent, the total bytes of application data sent, and the average flow duration (and therefore the average data rate).

@@

The simulations described in \Section{simulations} provide us with the raw data to compare performance metrics for synthetic traffic models with those for real traffic traces. To assess the realism of these models, however, we need a measure of how inaccurate the synthetic performance values are when compared to the real values. Let $x$ be the value of a performance metric using real traffic, and $y$ the value of the same metric using an alternate traffic model. Some common measures of error are the difference [$y-x$], the ratio [$y/x$], and the relative error [$(y-x)/x$], also commonly called the ``standard error.'' These are all reasonable measures of error; but which is most appropriate for assessing the realism of performance metrics? Instead of picking one arbitrarily, we will first consider the properties that an ideal error function should have, and then use those properties to determine the best measure of error. Moreover, we will show that the unique measure of error that exhibits these ideal properties is the  \textit{log-ratio} of metric values:
\begin{align}
\log(y/x)=\log(y)-\log(x).
\end{align}
In this discussion, $E(x,y)$ is a generic error function applied to the synthetic value, $y$, with respect to the real value, $x$.

The first property that an error function should have is  \textit{insensitivity to common factors}. That is, if both values are scaled by the same constant, the error should be unaffected:
\begin{align}
\label{eqn:normalization-invariance}
\forall\: x,y,c: E(x c,y c) &= E(x,y)
\end{align}
There are three major motivations for this requirement:
\begin{enumerate}
\item Changing units should not affect error values.
\item Error values for ``large'' and ``small'' scenarios should be directly comparable. Scenarios with large $x$ values will naturally have larger raw differences between $x$ and $y$. This requirement allows scenarios of different scales to be compared fairly and without bias.
\item Changing between metrics that differ by a known constant for each scenario should not affect error values.
\end{enumerate}
The last point is best illustrated by an example. Consider two closely related performance metrics: average throughput, $t$, and total bytes received, $r$. Suppose that there are $f$ flows in a given scenario with average duration, $d$. Since $t =r/fd$, the metrics $t$ and $r$ contain the same information---they differ only by a known constant in each simulation scenario. Equation~\ref{eqn:normalization-invariance} ensures that the errors of these metrics are the same:
\begin{align}
E(t_{\text{trc}},t_{\text{mod}})
	= E\left({\frac{r_{\text{trc}}}{fd},\frac{r_{\text{mod}}}{fd}}\right)
	= E(r_{\text{trc}},r_{\text{mod}}).
\end{align}
The difference measure does not satisfy Equation~\ref{eqn:normalization-invariance}, but the ratio, relative error, and log-ratio error measures all do.

@@

It is easily verified that the difference, ratio, and relative error measures do not satisfy \Equation{linear-compounding}, and the difference, as noted, does not satisfy \Equation{normalization-invariance}. The log-ratio is the only metric presented that satisfies both conditions. Moreover, it can be proved that $\log(y/x)$ is the \textit{only} differentiable function that satisfies both (up to a constant). In the Appendix, we present a proof of this claim. Throughout the rest of the paper, we use the log-ratio to measure the error of performance metrics.
% TODO: should Appendix be capitalized?
% TODO: hyperref the Appendix for PDF.

@@

\begin{sidewaysfigure*}
\subfloat[\textbf{Application:} Average End-to-End Delay]%
{\includegraphics[width=4.45in]{plots/Application/delay.pdf}}\hfill
\subfloat[\textbf{Application:} Packet Delivery Ratio]%
{\includegraphics[width=4.45in]{plots/Application/delivery_ratio.pdf}}
%
\subfloat[\textbf{Application:} Average Received Throughput]%
{\includegraphics[width=4.45in]{plots/Application/throughput.pdf}}\hfill
\subfloat[\textbf{Network:} AODV Control Overhead]%
{\includegraphics[width=4.45in]{plots/Network/control_packets.pdf}}
%
\subfloat[\textbf{Network:} Packets Dropped in Routing Queue]%
{\includegraphics[width=4.45in]{plots/Network/dropped.pdf}}\hfill
\subfloat[\textbf{Link:} 802.11 Control Overhead]%
{\includegraphics[width=4.45in]{plots/Link/overhead.pdf}}%
%
\caption{\footnotesize%
Box-and-whisker plots of log-ratio error values for all metrics and traffic models. The lower axis indicates the log-ratio, while the upper axis shows raw ratio values. Each box contains the central majority of log-ratio values: the left and right bounds are at the $25^\text{th}$ and $75^\text{th}$ percentiles. The dark middle line indicates the median value, while the diamond marks the mean. The whiskers (dotted lines) extend to the furthest non-outlier values, while the points beyond that are outliers. The notch in the middle of each bar indicates a $95\%$ confidence interval for the true underlying median value; if two notches do not overlap, they are very unlikely to have the same median.
}
\label{fig:box-plots}
\end{sidewaysfigure*}
